---
title: "Data Sources"
format:  
  html:
    code-fold: true
    embed-resources: true
editor_options: 
  chunk_output_type: inline
---

# U.S. Bureau of Transportation Statistics

The U.S. Bureau of Transportation Statistics (BTS) is a prominent agency within the United States Department of Transportation. Established to provide comprehensive and objective transportation statistics, the BTS plays a crucial role in collecting, analyzing, and disseminating data related to various modes of transportation. This includes air, water, rail, and road transport, as well as aspects like freight movement, passenger travel, and economic indicators within the transportation sector. By offering reliable and up-to-date information, the BTS serves as a valuable resource for policymakers, researchers, and the general public, contributing to informed decision-making and a deeper understanding of the nation's transportation landscape. You can access this data at [website](https://www.bts.gov/).

**Source**  [download data](https://data.bts.gov/Research-and-Statistics/Monthly-Transportation-Statistics/crem-w557/explore)

## Sector 1: Moving Goods in the United States

### Value of shipments by transportation mode
```{r}
df1 <- read.csv("data/clean_data/11value_us.csv")
#df1$Year <- as.Date(ISOdate(df1$Year, 1, 1))
head(df1)
```

### Weight of shipments by transportation mode
```{r}
df2 <- read.csv("data/clean_data/12weight_us.csv")
head(df2)
```

### Total freight moved by distance
```{r}
df3 <- read.csv("data/clean_data/13distance_us.csv")
head(df3)
```

## Sector 2: International Freight

### U.S.-international freight trade by transportation mode
```{r}
df4 <- read.csv("data/clean_data/21international_value.csv")
head(df4)
```

### Value of U.S.-international freight flows by region and transportation mode
```{r}
df5 <- read.csv("data/clean_data/22international_Region.csv")
head(df5)
```

## Sector 3: Contribution to the U.S. Economy

### Gross domestic product (GDP) attributed to transportation by mode
```{r}
df6 <- read.csv("data/clean_data/31GDP.csv")
head(df6)
```

### Monthly transportation employment by mode
```{r}
df7 <- read.csv("data/clean_data/32monthly_employment.csv")
head(df7)
```

### Average freight revenue per ton-mile
```{r}
df8 <- read.csv("data/clean_data/33revenue.csv")
head(df8)
```

### Monthly diesel and jet fuel prices
```{r}
df9 <- read.csv("data/clean_data/34monthly_fuel_prices.csv")
head(df9)
```

### Monthly Freight Transportation Services Index (TSI)
```{r}
df10 <- read.csv("data/clean_data/35monthly_TSI.csv")
head(df10)
```

# Yahoo Stocks Data

Yahoo! Finance, a media property within the Yahoo! network, is a comprehensive platform offering financial news, data, commentary, stock quotes, press releases, financial reports, and original content. Not only does it furnish online tools for personal finance management, but it also publishes both partner content and original stories by its in-house journalists. Holding the 20th position in SimilarWeb's major news and media websites ranking, Yahoo! Finance has been a prominent player in delivering financial information. In 2017, the platform expanded its coverage to include cryptocurrency news, providing insights into more than 9,000 unique coins, including popular ones like Bitcoin and Ethereum. [website](https://finance.yahoo.com/) 

## Sector 4: Changes in the Most Vital Stocks

**Source:** 

```{r}
#| warning: false

library(quantmod)
library(tidyquant)
library(ggplot2)
library(plotly)
library(ggfortify)
options("getSymbols.warning4.0"=FALSE)
options("getSymbols.yahoo.warning"=FALSE)

tickers = c("UPS","UNP","JBHT","KEX")
for (i in tickers){
  getSymbols(i,
             from = "2017-01-01",
             to = "2023-12-31")}

x <- list(
  title = "date"
)
y <- list(
  title = "value"
)

stock <- data.frame(UPS$UPS.Adjusted,
                    UNP$UNP.Adjusted,
                    JBHT$JBHT.Adjusted,
                    KEX$KEX.Adjusted
                    )


stock <- data.frame(stock,rownames(stock))
colnames(stock) <- append(tickers,'Dates')

stock$date<-as.Date(stock$Dates,"%Y-%m-%d")
(head(stock))
```


# World Health Organization: Coronavirus (COVID-19)
The COVID-19 dashboard by the World Health Organization (WHO) offers daily updates on official counts of COVID-19 cases, deaths, and vaccine utilization reported by various countries, territories, and areas. This interactive dashboard serves as a dynamic resource for data visualization, dissemination, and exploration, providing users with regularly updated information. Additionally, it facilitates access to other valuable and informative resources related to the COVID-19 pandemic.

**Source**  [COVID-19 data](https://github.com/owid/covid-19-data/tree/master/public/data)

**COVID-19 New Cases and Total Cases in the U.S.**

```{r}
covid = read.csv("./data/clean_data/covid.csv",header =TRUE, sep = ',')
covid$date <- as.Date(covid$date)
covid$new_cases_per_million <- as.double(covid$new_cases_per_million)
covid$total_cases_per_million <- as.double(covid$total_cases_per_million)
head(covid)
```

```{r}
par(mfrow = c(2, 1)) 
plot1 = plot(covid$date, covid$new_cases_per_million, col="darkblue", xlab="Time", ylab="New Cases(per Million)")
title("Covid New Cases in U.S. Over Time")

plot2 = plot(covid$date, covid$total_cases_per_million, col="lightblue", xlab="Time", ylab="Total Cases(per Million)")
title("Covid Total Cases in U.S. Over Time")
```
